Towards a Cognitive System That Can Recognize Spatial Regions Based on Context

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/evaluation.tex

We evaluated our approach on six classrooms (two simulated and four real) and two simulated studio apartments. The simulated rooms were based on real-life counterparts. For each room we manually encoded the appropriate CDSRs that could be represented by our approach. For the classrooms these were the front and back, and the front and back rows of desk. For the studios these were the kitchen, office and living areas. These manually encoded regions were used as the base CDSRs for analogical transfers, and can be considered the training data for our evaluation.

@@

We consider a \textit{problem instance} to be a room and a sought CDSR type. For each room for which we have a human-labelled target region of the sought type, we generate an \textit{inferred region} using analogical transfer. To assess the quality of the transfer, we calculate precision ($p$, the proportion of the inferred region that overlaps with the target region) and recall ($r$, the proportion of the target region that overlaps with the inferred region) as follows:

\begin{equation}
	p=\frac{area(inferred\ region \cap target\ region)}{area(inferred\ region)}
\end{equation}
\begin{equation}
	r=\frac{area(inferred\ region \cap target\ region)}{area(target\ region)}
\end{equation}

Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the manually annotated CDSR in the target and appropriate target region; and the region for the whole room and the target region. 
 Results comparing transferred and target regions measure how well the output of our system matches user expectations. Results comparing the manual annotations to the target regions measure how well the anchor point-based regions we chose match the users' regions (who were not constrained to anchor points). Results comparing whole rooms to target regions provide a baseline performance for comparison.

\subsection{Results}
\begin{table*}
	\center
\begin{tabular}{|c|c|c|}
\hline
Inferred & Manually Encoded & Entire Room \\
\hline
$\bar{p}$=.47 $\sigma$=.37, $\bar{r}$=.46 $\sigma$=.38 & $\bar{p}$=.71 $\sigma$=.30, $\bar{r}$=.67 $\sigma$=.25 & $\bar{p}$=.17 $\sigma$=.11, $\bar{r}$=.98 $\sigma$=.05  \\
\hline
\end{tabular}
\caption{Overall Performance}
  \label{tab:overall}

@@

To assess overall performance, Table~\ref{tab:overall} summarizes the results across all problem instances against user-defined target regions from three different users. The inferred regions achieved a precision of .47 ($\sigma$=.37) and a recall of .46 ($\sigma$=.38). Comparing the manually encoded regions against each target region results in a mean precision of .71 ($\sigma$=.30) and recall of .67 ($\sigma$=.25). The region defined by the room corresponds to the target region with a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).

@@

These results support the hypothesis that anchor points can provide a symbolic representation on top of sensor data for context-dependent spatial regions, and, when combined with qualitative spatial relations, they facilitate learning from a single example through analogical transfer. Collaboration with human users requires a high precision and recall, because cognitive systems must be able to understand as well as refer to these regions in human environments. Consequently, the high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation. Our future work seeks to further evaluate the utility of this representation by embedding the cognitive system within a tasks with human users.

>>>>

We evaluated our approach on six classrooms (two simulated and four real) and two simulated studio apartments. The simulated rooms were based on real-life counterparts. For each room we manually encoded the appropriate CDSRs that could be represented by our approach. For the classrooms these were the front and back, and the front and back rows of desks. For the studios these were the kitchen, office and living areas. These manually encoded regions were used as the base CDSRs for analogical transfers, and can be considered the training data for our evaluation.

@@

We consider a \textit{problem instance} to be a room and a sought CDSR type. For each room containing a manually encoded CDSR of the sought type, we generate an \textit{transferred region} using analogical transfer. To assess the quality of the transfer, we calculate precision ($p$, the proportion of the transferred region that overlaps with the target region) and recall ($r$, the proportion of the target region that overlaps with the transferred region) as follows:

\begin{equation}
	p=\frac{area(transferred\ region \cap target\ region)}{area(transferred\ region)}
\end{equation}
\begin{equation}
	r=\frac{area(transferred\ region \cap target\ region)}{area(transferred\ region)}
\end{equation}

Using this approach we generate results showing the matches between each of the following pairs of regions: the transferred region and the appropriate target region; the manually encoded CDSR and target region; and the region for the whole room and the target region. 
Results comparing transferred and target regions measure how well our system is able apply a single example to new situations. The comparisons between the manual annotations to the target regions measure how well the anchor points we chose capture the users' regions (who were not constrained to anchor points). Results from the whole room regions provide a baseline performance for comparison.

\subsection{Results}
\begin{table*}
	\center
\begin{tabular}{|c|c|c|}
\hline
Transferred & Manually Encoded & Entire Room \\
\hline
$\bar{p}$=.47 $\sigma$=.37, $\bar{r}$=.46 $\sigma$=.38 & $\bar{p}$=.71 $\sigma$=.30, $\bar{r}$=.67 $\sigma$=.25 & $\bar{p}$=.17 $\sigma$=.11, $\bar{r}$=.98 $\sigma$=.05  \\
\hline
\end{tabular}
\caption{Overall Performance}
  \label{tab:overall}

@@

To assess overall performance, Table~\ref{tab:overall} summarizes the results across all problem instances against user-defined target regions from three different users. The transferred regions achieved a precision of .47 ($\sigma$=.37) and a recall of .46 ($\sigma$=.38). Comparing the manually encoded regions against each target region results in a mean precision of .71 ($\sigma$=.30) and recall of .67 ($\sigma$=.25). The region defined by the room corresponds to the target region with a precision of .17 ($\sigma$=.11) and recall of .98 ($\sigma$=.05).

@@

These results support the hypothesis that anchor points can provide a symbolic representation on top of sensor data for context-dependent spatial regions, and, when combined with qualitative spatial relations, they facilitate learning from a single example through analogical transfer. Collaboration with human users requires a high precision and recall, because cognitive systems must be able to understand as well as refer to these regions in human environments. Consequently, the high manually encoded precisions and recalls indicate that the defined anchor points are a reasonable starting place for a symbolic representation. Our future work seeks to further evaluate the utility of this representation by embedding the cognitive system within tasks with human users.

### 2012-01-25 scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex

\fw{
(regionType CDSR9 FrontRegion) \\
(regionBoundarySegment CDSR9 \\
\hspace*{2em}(YMaxXFewestFn Room-Base) \\
\hspace*{2em}(YMinXFewestFn Room-Base)) \\
(regionBoundarySegment CDSR9 \\
\hspace*{2em}(YMinXFewestFn Room-Base) \\
\hspace*{2em}(YMinXFewestFn Group1))\\
...
}
}
  \caption{Representing the front of the classroom context-dependent region \fw{CDSR9}}
  \label{fig:cdsr-reps}
\end{figure}

Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room1} which is pictured in the top of Figure~\ref{fig:rooms}. The anchor points (connected in orange in Figure~\ref{fig:rooms}) define the region. \fw{(LeftmostBottomFn Room1)} and \fw{(LeftmostTopFn Room1)} are the points with the lowest x coordinate out of the set of points defined as the top and bottom of \fw{Room1} respectively. The sets of points representing the top and bottom of the classroom are found by selecting the points defining the area within a threshold of the area's maximum or minimum y coordinate. The other corners of the context-dependent spatial region are defined by the four desks and the whiteboard. Because perceived objects (e.g. the desks) are defined as single points, they can serve as anchor points directly. The semantic label \fw{FrontRegion} ties the geometric extent of the boundary points to a conceptual region. This definition for the front of the classroom is specific to \fw{Room1} and its entities. It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary. If the desks were in a different position then the region would cover a different extent (e.g. if they were closer to the whiteboard then the region would be smaller).

@@

% \from{Nick}{This section feels a bit long. Can we cut it down at all? Perhaps the second half of the second paragraph if we don't use the constraints explicitly}
% done

We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding. To avoid burdening potential tutors with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training. For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building. We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic as well as geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.

Analogy is an essential cognitive process. In humans, analogical processing has been observed in language comprehension, problem-solving, and generalization \cite{Gentner2003}. The structure-mapping theory of analogy and similarity postulates this process as an alignment between two structured representations, a \textit{base} and a \textit{target} \cite{Gentner1983a}. 
% This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}. The identicality constraint provides a strong preference for only allowing identical predicates to match. Parallel connectivity states that if two predicates are matched then their arguments must also match. The one-to-one mapping constraint requires that each element in the base corresponds to at most one element in the target and vice versa. To select between competing mappings, the \textit{systematicity} principle prefers mappings that are highly interconnected and contain deep chains of higher order relations over mappings with an equal number of relations which are independent from each other.
We use the Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} to perform analogical matching in our system. Given base and target representations as input, SME produces one or more mappings. Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target structures.  Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences. SME operates in polynomial time, using a greedy algorithm \cite{Forbus/etal1994}.

@@

Figure \ref{fig:analogy} illustrates a sample mapping between six base expressions and three target ones. Each oval represents a binary predicate, and its two entity arguments are represented by squares. SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case . This is done in the following manner. First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME. Additional align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}. While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa. In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows. Next, SME creates a candidate inference for the region boundary expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping. The candidate inference is stated as \fw{(regionBoundary(:skolem CSDR1) (XMinYMost ...))}. \fw{:skolem} is used to denote entities that do not exist in the target description but are included in the candidate inferences. In this case, the candidate inference states that there is something like \fw{CSDR1} in the target, and its boundary includes the \fw{XMinYMost} point of the two desk group. Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions. %Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.

@@

We first create the base and target representations by running Dora in the two different classrooms. These are pictured in the top and bottom of Figure~\ref{fig:rooms} respectively. In each case, Dora is manually driven around the room to allow it to create a metric map. Once the map is created, Dora is then positioned such that the objects are observable and the visual recognition system is run. The map and object data that results from this is then passed on to the QSR generator. In the base case Dora perceives 2 desks, a whiteboard and the room area. To this we add the CDSR representing the front of the room. The case includes a total of \textbf{SOMENUMBER} expressions relating the \textbf{SOMENUMBER} entities. \textbf{SOMENUMBER} of these expressions are used to define the boundary of a CDSR representing the front of the room: the \textbf{SOMENUMBER} desks in the front row, the whiteboard, and the leftmost bottom and leftmost top of the room. The target case includes \textbf{SOMENUMBER}. The target case includes \textbf{SOMENUMBER} statements and \textbf{SOMENUMBER} entities.

SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation. The resulting analogy includes \textbf{SOMENUMBER} correspondences between the entities and expressions in the base and target and \textbf{SOMENUMBER} candidate inferences.  All \textbf{SOMENUMBER} \fw{regionBoundary} statements in the base appear as candidate inferences. \textbf{SOMENUMBER} of the anchor points are exactly correct, the \textbf{\dots}. The other \textbf{SOMENUMBER} analogically transferred anchor points are to the \textbf{\dots}. The green region in the lower image of Figure \ref{fig:rooms} illustrates the CDSR created by the \textbf{SOMENUMBER} boundary points in the target environment.

@@

In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions. The system identifies CDSRs in previously unseen environments through analogy with a manually annotated known room. The system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy. In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with future work. These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g. MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g. comparing the QSRs present in both base and transferred regions) to improve results. Despite these limitations our evaluation demonstrated that the system is able to match the CDSRs produced by trial subjects on unseen rooms to an acceptable degree \textbf{SOMENUMBER}.

>>>>

\fw{(regionType CDSR9 FrontRegion) \\
(boundarySegment CDSR9 \\
\hspace*{2em}(YMaxXFewestFn Room3) \\
\hspace*{2em}(YMinXFewestFn Room3)) \\
(boundarySegment CDSR9 \\
\hspace*{2em}(YMinXFewestFn Room3) \\
\hspace*{2em}(YMinXFewestFn Group1))
}
}
  \caption{Three of the five expressions representing the front of the classroom context-dependent region \fw{CDSR9}}
  \label{fig:cdsr-reps}
\end{figure}

Figure \ref{fig:cdsr-reps} contains three of the five expressions defining the front of classroom \fw{Room3} which is pictured in the top of Figure~\ref{fig:rooms}. The boundary segments (shown in orange in Figure~\ref{fig:rooms}) define the extent of the region. \fw{(YMaxXFewestFn Room3)} and \fw{(YMinXFewestFn Room3)} are the points with the highest and lowest y coordinate out of the set of points within 5\% of the minimum x coordinate of \fw{Room3}. The next segment connect the lower left coordinate in the figure to the \fw{(YMinXFewestFn Group1)}, where \fw{Group1} includes the eight desks. There are two more boundary segments completing a polygon for this region. The semantic label \fw{FrontRegion} ties this polygon to a conceptual region, "the front of the room". This definition for the front of the room is specific to \fw{Room3} and its entities. It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary. If the desks were in a different position then the region would cover a different extent (e.g. if they were further to the left then the region would be smaller).

@@

We assume that a cognitive system will have a way of initially acquiring examples of CDSRs, e.g., by being taught through dialogue, sketching, or hand-coding. To avoid burdening potential users with the task of teaching the system every CDSR individually, it is desirable for a cognitive system to be able to automatically recognize similar regions after initial training. For example, after a janitorial robot has been taught where the front of one classroom is, it should be able to identify the fronts of other classrooms in the building. We have chosen \emph{analogy} as the approach our system will use to solve this problem, as analogy has been previously used to successfully combine semantic and geometric information in spatial reasoning tasks~\cite{Lockwood/etal2008}.

Analogy is an essential cognitive process. In humans, analogical processing has been observed in language comprehension, problem-solving, and generalization \cite{Gentner2003}. The structure-mapping theory of analogy and similarity postulates this process as an alignment between two structured representations, a \textit{base} and a \textit{target} \cite{Gentner1983a}. 
% This alignment process is governed by three constraints: \textit{identicality}, \textit{parallel connectivity}, and \textit{one-to-one mapping}. The identicality constraint provides a strong preference for only allowing identical predicates to match. Parallel connectivity states that if two predicates are matched then their arguments must also match. The one-to-one mapping constraint requires that each element in the base corresponds to at most one element in the target and vice versa. To select between competing mappings, the \textit{systematicity} principle prefers mappings that are highly interconnected and contain deep chains of higher order relations over mappings with an equal number of relations which are independent from each other.
We use the Structure-Mapping Engine (SME) \cite{Falkenhainer1989a} to perform analogical matching in our system. Given base and target representations as input, SME produces one or more mappings. Each mapping is represented by a set of \textit{correspondences} between entities and expressions in the base and target structures. Mappings are defined by expressions with an identical relation and corresponding arguments. When provided with expression strengths, such as, our spatial relationships, SME prefers mappings with closely aligned fact strengths. SME can be given \textit{pragmatic constraints} that require certain entities in the base to be included in the mapping. Mappings also include \textit{candidate inferences} which are conjectures about the target using expressions from the base which, while unmapped in their entirety, have subcomponents that participate in the mapping's correspondences. SME operates in polynomial time, using a greedy algorithm \cite{Forbus/etal1994}.

@@

Figure \ref{fig:analogy} illustrates a sample mapping between six base expressions and three target ones. Each oval represents a predicate, and the entity arguments are represented by squares. SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case in the following manner. First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME. Then SME aligns the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}. While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa. In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows. Next, SME creates a candidate inference for the boundary segment expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping. The candidate inference is shown in red in the figure. Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions. %Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.

@@

We first create the base and target representations by running Dora in the two different classrooms. These are pictured in the top and bottom of Figure~\ref{fig:rooms} respectively. In each case, Dora is manually driven around the room to allow it to create a metric map. Once the map is created, Dora is then positioned such that the objects are observable and the visual recognition system is run. The map and object data that result from this are then passed on to the QSR generator. In the base case, Dora perceives 8 individual desks, a group entity containing these desks and the room area. To this we add the CDSR representing the front of the room. The case includes a total of 50 expression relating the 20 entities. Six of these expressions are used to define the boundary segments and CDSR representing the front of the room. The target case includes 26 expressions and 11 entities.

SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation requiring \fw{Room3} and \fw{Group1} participate in the mapping as they are referenced by the anchor points in the base. The resulting analogy includes 26 correspondences between the entities and expressions and 32 candidate inferences. Four of these candidate inferences define the CDSR in the target with anchor points to the group of desks and the room in the target. The green region in the lower image of Figure \ref{fig:rooms} illustrates the transferred CDSR.

@@

In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions. The system identifies CDSRs in previously unseen environments through analogy with a single example. This is a difficult cognitive systems task requiring integration of semantic and geometric knowledge to identify regions which can be as small as 8\% of the room. Our system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy to achieve this task. In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with future work. These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g. MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g. comparing the QSRs present in both base and transferred regions) to improve results. In addition, we must complement our system development work with more comprehensive human studies assessing how people define and use these regions as well as how well anchor points capture them. Despite the preliminary nature of this work, our evaluation demonstrates that the system is able to produce CDSRs, which overlap with user-defined region, from a single example for 6 out of 7 region types.
